---
title: "ONC ERDDAP with R"
---

Demonstration of how to use the `rerddap` package to access data from the 
Ocean Networks Canada ERDDAP server. A minimal translation from <https://github.com/oceanhackweek/ohw-tutorials/blob/OHW26/01-Tue/DataAccess/PartI_ERDDAP_Tutorial.ipynb>

This uses the [rerddap](https://docs.ropensci.org/rerddap/index.html) package to 
make the querying of ERDDAP servers easier. 

First, load the packages needed to query the ERDDAP server, wrangle the results, and combine plots.

```{r}
#| include: false
options(tidyverse.quiet = TRUE)
```

```{r}
#| label: load-packages
#| message: false
library(rerddap)
library(tidyverse)
library(patchwork)
```

Set up the query parameters: how far back in time to pull data, which
variables to request, and which dataset/server to query.

We are pulling the data from this record at China Creek: <https://dap.oceannetworks.ca/erddap/tabledap/scalar_1217325.html>

```{r}
base_url <- 'https://dap.oceannetworks.ca/erddap'
dataset_id <- 'scalar_1217325'

time_series_length <- 14 # in days. note ERDDAP server is slow, which is why we are limiting to 14 days for now.
start_time <- Sys.time() - time_series_length * 24 * 3600
query_vars <- c(
  'time',
  'temperature',
  'oxygen_corrected',
  'oxygen_uncorrected',
  'latitude',
  'longitude',
  'depth'
)
```

Inspect the dataset's metadata (available variables, units, coverage) before
querying it.

```{r}
info(dataset_id, url = base_url)
```

Pull the data for the last `time_series_length` days, restricted to the
fields we're interested in. Note the time constraint is passed as a single
unnamed string (`"time>=..."`), as required by `tabledap()`.

```{r}
cc_02_data <- tabledap(
  info(dataset_id, url = base_url),
  paste0("time>=", format(start_time, "%Y-%m-%dT%H:%M:%SZ")),
  fields = query_vars
)
```

Plot temperature and corrected oxygen over time, stacked together for
comparison.

```{r}
temp_plot <- cc_02_data |>
  ggplot(aes(x = time, y = temperature)) +
  geom_line(colour = "red") +
  labs(
    title = paste0("Temperature at depth ", unique(cc_02_data$depth), "m"),
    x = "Time",
    y = "Temperature (°C)"
  ) +
  theme_minimal()

oxy_plot <- cc_02_data |>
  ggplot(aes(x = time, y = oxygen_corrected)) +
  geom_line(colour = "blue") +
  labs(
    title = paste0("Corrected Oxygen at depth ", unique(cc_02_data$depth), "m"),
    x = "Time",
    y = "Oxygen (µmol/kg)"
  ) +
  theme_minimal()

temp_plot / oxy_plot
```
